### 2018-05-29 submission/manuscript.Rmd
update second section of results

loadLibs(c("tidyverse", "scales", "knitr", "rmarkdown", "stringr"))

@@

# Vector of sub samples used
sub_sample_level <- c("1000", "5000", "10000")

dist_data <- sapply(sub_sample_level, 
                    function(x) as.data.frame(read_csv(paste("../data/process/tables/bray_5_cycle_dist_",                                                x, "_data.csv", sep = ""))), simplify = F)

mock_dist_data <- sapply(sub_sample_level, 
                    function(x) as.data.frame(read_csv(paste("../data/process/tables/mock_bray_5_cycle_dist_", 
                                               x, "_data.csv", sep = ""))), simplify = F)

@@

***The Number of OTUs are Dependent on Polymerase:***
Differences in the range of the number of OTUs detected for fecal samples is dependent on the polymerase used (e.g. Accuprime at 35 cycles = `r filter(fecal_numOTU_data, cycles == "35x", taq == "Accuprime") %>% pull(taq_min)` - `r filter(fecal_numOTU_data, cycles == "35x", taq == "Accuprime") %>% pull(taq_max)` versus Phusion at 35 cycles = `r filter(fecal_numOTU_data, cycles == "35x", taq == "Phusion") %>% pull(taq_min)` - `r filter(fecal_numOTU_data, cycles == "35x", taq == "Phusion") %>% pull(taq_max)`) [Figure 1]. Additionally, there is a trend for lower number of cycles (15-20) to result in a reduced range in the number of OTUs versus higher number of cycles (25, 30, and 35) for all polymerases (e.g. Phusion at 15 cycles = `r filter(fecal_numOTU_data, cycles == "15x", taq == "Phusion") %>% pull(taq_min)` - `r filter(fecal_numOTU_data, cycles == "15x", taq == "Phusion") %>% pull(taq_max)` versus Phusion at 35 cycles = `r filter(fecal_numOTU_data, cycles == "35x", taq == "Phusion") %>% pull(taq_min)` - `r filter(fecal_numOTU_data, cycles == "35x", taq == "Phusion") %>% pull(taq_max)`) [Figure 1]. There is an overall difference in the number of OTUs detected within fecal samples between polymerases at 35 cycles (F-stat > `r round(min(filter(fs_anova_summary, cycle == "35x", sub_sample_level == 1000)[, "F.value"]), digits = 2)`, P-value = `r format(filter(fs_anova_summary, cycle == "35x", sub_sample_level == 1000)[, "bh"], digits = 2, nsmall = 2, scientific = T)`) [Table S1]. Using a Tukey post-hoc test to identify which polymerase groups were different from each other at 35 cycles, no difference was found (P-value `r ifelse(min(filter(fs_tukey_summary, sub_sample_level == 1000)[, "padj"]) > 0.05,  invisible(">"), invisible("check data"))` 0.05) [Table S2]. The polymerase dependent difference in the range of the number of OTUs also was observed in the mock community samples [Figure 2]. The closest the polymerases came to the total of `r pull(ref_mock_otu_num, numOtus)` OTUs created by the mock reference 16S sequences was at 25 and 30 cycles [Figure 2]. Regardless of if fecal or mock communities were used, the same polymerases generated high and low number of OTUs and this was consistent across the number of cycles used [Figure 1-2 & Table S1-S2].  In contrast to the results obtained with fecal samples, differences between polymerases for the number of OTUs created were observed as early as 20 cycles in the mock community (F-stat = `r round(filter(mock_anova_summary, sub_sample_level == 1000, cycle == "20x")[, "F.value"], digits = 2)`, P-value = `r format(filter(mock_anova_summary, sub_sample_level == 1000, cycle == "20x")[, "bh"], digits = 1, nsmall = 1)`) [Table S1]. Using a Tukey post-hoc test, the majority of differences for the number of OTUs detected in the mock community is largely due to Kappa and Platinum differences versus the other polymerases across the different number of cycles [Table S2]. Based on these observations in fecal and mock communities, it is clear that using different polymerases result in a different total number of OTUs within a sample.

***Minimal Bray-Curtis Differences are Detected and are Dependent on Cycle Number:***
A few small differences based on the number of cycles were detected in overall bacterial community composition. Within the same fecal sample and independent of polymerases, there are large differences in the community composition between 20 and 25 cycles [Figure 3A]. Additionally, when data was available for the mock communities, there are also large observed differences between 20 and 25 cycles [Figure 3B]. These community differences within a sample disappear when comparing 25 to 30 cycles and do not persist past 25 cycles [Figure 3]. Although these trends are clearly noticeable, there is no detectable difference in Bray-Curtis index when comparing to the previous 5-cycle increment for both fecal and mock communities after multiple comparison correction (P-value `r ifelse(min(filter(bray_sim_to_prev_cycle_kruskal_results, sub_sample_level == 1000)[, "bh"], na.rm = T) > 0.05, invisible(">"), invisible("check data"))` 0.05). Using PERMANOVA to test for community differences based on the number of cycles within polymerases, only Phusion had cycle dependent differences (P-value = `r format(filter(bray_permanova_by_taq_results, sub_sample_level == 1000, taq == "PHU")[, "bh"], digits = 1, nsmall = 1)`. For fecal samples, Phusion was one of two polymerases that had enough sequences to be rarefied to 1000 at 15 cycles. Overall, these data suggest that there are small polymerase differences in Bray-Curtis index that are dependent on the number of cycles.

@@

**Figure 1: Total Number of OTUs in Fecal Samples by Number of Cycles.** The x-axis represents the different number of cycles used and the y-axis is the number of OTUs. The points represent the median value of all four fecal samples. The lines represent the range of the minimum and maximum number of OTUs detected within the four fecal samples. The range in the number of OTUs detected in the different fecal samples increased as cycle number increased. This increased range also was larger for specific HiFi DNA polymerases.

**Figure 2:  Total Number of OTUs in Mock Samples by Number of Cycles.** The x-axis represents the different number of cycles used and the y-axis is the number of OTUs. The points represent the median value of all four fecal samples. The lines represent the range of the minimum and maximum number of OTUs detected within the four fecal samples. The dotted black line represents the number of OTUs detected when only the references sequences for the mock community are clustered. The range in the number of OTUs detected in the mock samples increased as cycle number increased. This range was also larger for specific HiFi DNA polymerases.

**Figure 3: Bray-Curtis Community Differences by Five-Cycle Intervals.** A) within person differences based on the next 5-cycle PCR interval in fecal samples. B) Within replicate difference based on the next 5-cycle PCR interval in Mock samples.

>>>>

loadLibs(c("tidyverse", "scales", "knitr", "rmarkdown", "stringr", "FSA"))

@@

# Vector of sub samples used
dist_data <- read_csv("../data/process/tables/bray_5_cycle_dist_1000_data.csv") %>% 
  mutate(taq = factor(taq, 
                      levels = c("ACC", "K", "PHU", "PL", "Q5"), 
                      labels = c("Accuprime", "Kappa", "Phusion", "Platinum", "Q5"))) %>% 
  group_by(taq, cycle_compare) %>% 
  summarise(dist_median = median(distance, na.rm = T), 
            dist_min = min(distance, na.rm = T), 
            dist_max = max(distance, na.rm = T)) %>% 
  mutate(range_diff = dist_max - dist_min) %>% 
  ungroup() %>% 
  filter(cycle_compare != "25to35")

fecal_kruskal_test_by_cycle_compare_summary <- dist_data %>% 
  nest() %>% 
  mutate(kruskal_analysis = map(data, ~kruskal.test(dist_median ~ factor(cycle_compare), data = .x)), 
         summary_data = map(kruskal_analysis, broom::tidy)) %>% 
  select(summary_data) %>% unnest(summary_data)

dunns_test_by_cycle_compare_summary <- dist_data %>% 
  nest() %>% 
  mutate(dunn_analysis = map(data, ~dunnTest(dist_median ~ factor(cycle_compare), data = .x, method = "bh")[["res"]])) %>% 
  select(dunn_analysis) %>% unnest() %>% 
  arrange(Comparison)

mock_dist_data <- read_csv("../data/process/tables/mock_bray_5_cycle_dist_1000_data.csv") %>% 
  mutate(taq = factor(taq, 
                      levels = c("ACC", "K", "PHU", "PL", "Q5"), 
                      labels = c("Accuprime", "Kappa", "Phusion", "Platinum", "Q5"))) %>% 
  filter(!is.na(distance)) %>% 
  group_by(taq, cycle_compare) %>% 
  summarise(dist_median = median(distance, na.rm = T), 
            dist_min = min(distance, na.rm = T), 
            dist_max = max(distance, na.rm = T)) %>% 
  mutate(range_diff = dist_max - dist_min) %>% 
  ungroup() %>% 
  filter(cycle_compare != "25to35")

mock_kruskal_test_by_cycle_compare_summary <- mock_dist_data %>% 
  nest() %>% 
  mutate(kruskal_analysis = map(data, ~kruskal.test(dist_median ~ factor(cycle_compare), data = .x)), 
         summary_data = map(kruskal_analysis, broom::tidy)) %>% 
  select(summary_data) %>% unnest(summary_data)

@@

***The number of OTUs are dependent on polymerase:***
Differences in the range of the number of OTUs detected for fecal samples is dependent on the polymerase used (e.g. Accuprime at 35 cycles = `r filter(fecal_numOTU_data, cycles == "35x", taq == "Accuprime") %>% pull(taq_min)` - `r filter(fecal_numOTU_data, cycles == "35x", taq == "Accuprime") %>% pull(taq_max)` versus Phusion at 35 cycles = `r filter(fecal_numOTU_data, cycles == "35x", taq == "Phusion") %>% pull(taq_min)` - `r filter(fecal_numOTU_data, cycles == "35x", taq == "Phusion") %>% pull(taq_max)`) [Figure 1]. Additionally, there is a trend for lower number of cycles (15-20) to result in a reduced range in the number of OTUs versus higher number of cycles (25, 30, and 35) for all polymerases (e.g. Phusion at 15 cycles = `r filter(fecal_numOTU_data, cycles == "15x", taq == "Phusion") %>% pull(taq_min)` - `r filter(fecal_numOTU_data, cycles == "15x", taq == "Phusion") %>% pull(taq_max)` versus Phusion at 35 cycles = `r filter(fecal_numOTU_data, cycles == "35x", taq == "Phusion") %>% pull(taq_min)` - `r filter(fecal_numOTU_data, cycles == "35x", taq == "Phusion") %>% pull(taq_max)`) [Figure 1]. There is an overall difference in the number of OTUs detected within fecal samples between polymerases at 35 cycles (F-stat > `r round(min(filter(fs_anova_summary, cycle == "35x", sub_sample_level == 1000)[, "F.value"]), digits = 2)`, P-value = `r format(filter(fs_anova_summary, cycle == "35x", sub_sample_level == 1000)[, "bh"], digits = 2, nsmall = 2, scientific = T)`) [Table S1]. Using a Tukey post-hoc test to identify which polymerase groups were different from each other at 35 cycles, no difference was found (P-value `r ifelse(min(filter(fs_tukey_summary, sub_sample_level == 1000)[, "padj"]) > 0.05,  invisible(">"), invisible("check data"))` 0.05) [Table S2]. The polymerase dependent difference in the range of the number of OTUs also was observed in the mock community samples [Figure 2]. The closest the polymerases came to the total of `r pull(ref_mock_otu_num, numOtus)` OTUs created by the mock reference 16S sequences was at 25 and 30 cycles [Figure 2]. Regardless of if fecal or mock communities were used, the same polymerases generated high and low number of OTUs and this was consistent across the number of cycles used [Figure 1-2 & Table S1-S2].  In contrast to the results obtained with fecal samples, differences between polymerases for the number of OTUs created were observed as early as 20 cycles in the mock community (F-stat = `r round(filter(mock_anova_summary, sub_sample_level == 1000, cycle == "20x")[, "F.value"], digits = 2)`, P-value = `r format(filter(mock_anova_summary, sub_sample_level == 1000, cycle == "20x")[, "bh"], digits = 1, nsmall = 1)`) [Table S1]. Using a Tukey post-hoc test, the majority of differences for the number of OTUs detected in the mock community is largely due to Kappa and Platinum differences versus the other polymerases across the different number of cycles [Table S2]. Based on these observations in fecal and mock communities, it is clear that using different polymerases result in a different total number of OTUs within a sample.

***The bacterial community is similar within polymerase and varies by number of cycles:***
Within each respective 5-cycle increment comparison there was no difference in Bray-Curtis index between the polymerases (P-value `r ifelse(min(filter(bray_sim_to_prev_cycle_kruskal_results, sub_sample_level == 1000)[, "bh"], na.rm = T) > 0.05, invisible(">"), invisible("check data"))` 0.05) [Figure 3]. Using PERMANOVA to test for community differences based on any of the number of cycles within polymerases, only Phusion had cycle dependent differences (P-value = `r format(filter(bray_permanova_by_taq_results, sub_sample_level == 1000, taq == "PHU")[, "bh"], digits = 1, nsmall = 1)`. For fecal samples, Phusion was one of two polymerases that had enough sequences to be rarefied to 1000 at 15 cycles. There was an overall decrease in Bray-Curtis index by cycle comparison group (e.g. the 15 cycle versus 20 cycle group compared to the 30 cycle to 35 cycle group) (P-value `r ifelse(pull(fecal_kruskal_test_by_cycle_compare_summary, p.value) < 0.01, invisible("<"), invisible(">"))` 0.01) [Figure 3A]. Using a Dunn's post-hoc test the 15 cycle versus 20 cycle and 20 cycle versus 25 cycle comparison groups had a higher Bray-Curits index than the 30 cycle versus 35 cycle comparison group `r ifelse(min(pull(dunns_test_by_cycle_compare_summary, P.adj)) < 0.05, invisible("<"), invisible(">"))` 0.05). The mock community has similar trends to the obeservations reported for the fecal bacterial community but none were significant `r ifelse(pull(mock_kruskal_test_by_cycle_compare_summary, p.value) < 0.05, invisible("<"), invisible(">"))` 0.05) [Figure 3B]. Overall, these data suggest that the number of cycles can change the bacterial community independent of the polymerase used to generate the sequences.

@@

**Figure 1: Total Number of OTUs in Fecal Samples by Number of Cycles.** The points represent the median number of OTUs of all four fecal samples. The lines represent the range of the minimum and maximum number of OTUs detected within the four fecal samples. The range in the number of OTUs detected in the different fecal samples increased as cycle number increased. This increased range also was larger for specific HiFi DNA polymerases.

**Figure 2:  Total Number of OTUs in Mock Samples by Number of Cycles.** The points represent the median number of OTUs for the mock samples. The lines represent the range of the minimum and maximum number of OTUs detected within the four fecal samples. The dotted black line represents the number of OTUs detected when only the references sequences for the mock community are clustered. The range in the number of OTUs detected in the mock samples increased as cycle number increased. This range was also larger for specific HiFi DNA polymerases.

**Figure 3: Bray-Curtis Community Differences by Five-Cycle Intervals.** A) within person differences based on the next 5-cycle PCR interval in fecal samples. B) Within replicate difference based on the next 5-cycle PCR interval in mock samples. The points represent the median Bray-Curtis index for the samples. The lines represent the range of the minimum and maximum Bray-Curits index value for each PCR 5-cycle increment comparison. The closer a sample is to a Bray-Curtis index of 1.00 the more dissimilar the bacterial community is of the two compared number of cycles.

### 2018-05-29 submission/manuscript.tex
update second section of results

\textbf{\emph{The Number of OTUs are Dependent on Polymerase:}}
Differences in the range of the number of OTUs detected for fecal
samples is dependent on the polymerase used (e.g.~Accuprime at 35 cycles
= 84 - 106 versus Phusion at 35 cycles = 84 - 136) {[}Figure 1{]}.
Additionally, there is a trend for lower number of cycles (15-20) to
result in a reduced range in the number of OTUs versus higher number of
cycles (25, 30, and 35) for all polymerases (e.g.~Phusion at 15 cycles =
10 - 19 versus Phusion at 35 cycles = 84 - 136) {[}Figure 1{]}. There is
an overall difference in the number of OTUs detected within fecal
samples between polymerases at 35 cycles (F-stat \textgreater{} 16.35,
P-value = 9.7e-05) {[}Table S1{]}. Using a Tukey post-hoc test to
identify which polymerase groups were different from each other at 35
cycles, no difference was found (P-value \textgreater{} 0.05) {[}Table
S2{]}. The polymerase dependent difference in the range of the number of
OTUs also was observed in the mock community samples {[}Figure 2{]}. The
closest the polymerases came to the total of 8 OTUs created by the mock
reference 16S sequences was at 25 and 30 cycles {[}Figure 2{]}.
Regardless of if fecal or mock communities were used, the same
polymerases generated high and low number of OTUs and this was
consistent across the number of cycles used {[}Figure 1-2 \& Table
S1-S2{]}. In contrast to the results obtained with fecal samples,
differences between polymerases for the number of OTUs created were
observed as early as 20 cycles in the mock community (F-stat = 15.82,
P-value = 0.002) {[}Table S1{]}. Using a Tukey post-hoc test, the
majority of differences for the number of OTUs detected in the mock
community is largely due to Kappa and Platinum differences versus the
other polymerases across the different number of cycles {[}Table S2{]}.
Based on these observations in fecal and mock communities, it is clear
that using different polymerases result in a different total number of
OTUs within a sample.

\textbf{\emph{Minimal Bray-Curtis Differences are Detected and are
Dependent on Cycle Number:}} A few small differences based on the number
of cycles were detected in overall bacterial community composition.
Within the same fecal sample and independent of polymerases, there are
large differences in the community composition between 20 and 25 cycles
{[}Figure 3A{]}. Additionally, when data was available for the mock
communities, there are also large observed differences between 20 and 25
cycles {[}Figure 3B{]}. These community differences within a sample
disappear when comparing 25 to 30 cycles and do not persist past 25
cycles {[}Figure 3{]}. Although these trends are clearly noticeable,
there is no detectable difference in Bray-Curtis index when comparing to
the previous 5-cycle increment for both fecal and mock communities after
multiple comparison correction (P-value \textgreater{} 0.05). Using
PERMANOVA to test for community differences based on the number of
cycles within polymerases, only Phusion had cycle dependent differences
(P-value = 0.03. For fecal samples, Phusion was one of two polymerases
that had enough sequences to be rarefied to 1000 at 15 cycles. Overall,
these data suggest that there are small polymerase differences in
Bray-Curtis index that are dependent on the number of cycles.

@@

\textbf{Figure 1: Total Number of OTUs in Fecal Samples by Number of
Cycles.} The x-axis represents the different number of cycles used and
the y-axis is the number of OTUs. The points represent the median value
of all four fecal samples. The lines represent the range of the minimum
and maximum number of OTUs detected within the four fecal samples. The
range in the number of OTUs detected in the different fecal samples
increased as cycle number increased. This increased range also was
larger for specific HiFi DNA polymerases.

\textbf{Figure 2: Total Number of OTUs in Mock Samples by Number of
Cycles.} The x-axis represents the different number of cycles used and
the y-axis is the number of OTUs. The points represent the median value
of all four fecal samples. The lines represent the range of the minimum
and maximum number of OTUs detected within the four fecal samples. The
dotted black line represents the number of OTUs detected when only the
references sequences for the mock community are clustered. The range in
the number of OTUs detected in the mock samples increased as cycle
number increased. This range was also larger for specific HiFi DNA
polymerases.

\textbf{Figure 3: Bray-Curtis Community Differences by Five-Cycle
Intervals.} A) within person differences based on the next 5-cycle PCR
interval in fecal samples. B) Within replicate difference based on the
next 5-cycle PCR interval in Mock samples.

>>>>

\textbf{\emph{The number of OTUs are dependent on polymerase:}}
Differences in the range of the number of OTUs detected for fecal
samples is dependent on the polymerase used (e.g.~Accuprime at 35 cycles
= 84 - 106 versus Phusion at 35 cycles = 84 - 136) {[}Figure 1{]}.
Additionally, there is a trend for lower number of cycles (15-20) to
result in a reduced range in the number of OTUs versus higher number of
cycles (25, 30, and 35) for all polymerases (e.g.~Phusion at 15 cycles =
10 - 19 versus Phusion at 35 cycles = 84 - 136) {[}Figure 1{]}. There is
an overall difference in the number of OTUs detected within fecal
samples between polymerases at 35 cycles (F-stat \textgreater{} 16.35,
P-value = 9.7e-05) {[}Table S1{]}. Using a Tukey post-hoc test to
identify which polymerase groups were different from each other at 35
cycles, no difference was found (P-value \textgreater{} 0.05) {[}Table
S2{]}. The polymerase dependent difference in the range of the number of
OTUs also was observed in the mock community samples {[}Figure 2{]}. The
closest the polymerases came to the total of 8 OTUs created by the mock
reference 16S sequences was at 25 and 30 cycles {[}Figure 2{]}.
Regardless of if fecal or mock communities were used, the same
polymerases generated high and low number of OTUs and this was
consistent across the number of cycles used {[}Figure 1-2 \& Table
S1-S2{]}. In contrast to the results obtained with fecal samples,
differences between polymerases for the number of OTUs created were
observed as early as 20 cycles in the mock community (F-stat = 15.82,
P-value = 0.002) {[}Table S1{]}. Using a Tukey post-hoc test, the
majority of differences for the number of OTUs detected in the mock
community is largely due to Kappa and Platinum differences versus the
other polymerases across the different number of cycles {[}Table S2{]}.
Based on these observations in fecal and mock communities, it is clear
that using different polymerases result in a different total number of
OTUs within a sample.

\textbf{\emph{The bacterial community is similar within polymerase and
varies by number of cycles:}} Within each respective 5-cycle increment
comparison there was no difference in Bray-Curtis index between the
polymerases (P-value \textgreater{} 0.05) {[}Figure 3{]}. Using
PERMANOVA to test for community differences based on any of the number
of cycles within polymerases, only Phusion had cycle dependent
differences (P-value = 0.03. For fecal samples, Phusion was one of two
polymerases that had enough sequences to be rarefied to 1000 at 15
cycles. There was an overall decrease in Bray-Curtis index by cycle
comparison group (e.g.~the 15 cycle versus 20 cycle group compared to
the 30 cycle to 35 cycle group) (P-value \textless{} 0.01) {[}Figure
3A{]}. Using a Dunn's post-hoc test the 15 cycle versus 20 cycle and 20
cycle versus 25 cycle comparison groups had a higher Bray-Curits index
than the 30 cycle versus 35 cycle comparison group \textless{} 0.05).
The mock community has similar trends to the obeservations reported for
the fecal bacterial community but none were significant \textgreater{}
0.05) {[}Figure 3B{]}. Overall, these data suggest that the number of
cycles can change the bacterial community independent of the polymerase
used to generate the sequences.

@@

\textbf{Figure 1: Total Number of OTUs in Fecal Samples by Number of
Cycles.} The points represent the median number of OTUs of all four
fecal samples. The lines represent the range of the minimum and maximum
number of OTUs detected within the four fecal samples. The range in the
number of OTUs detected in the different fecal samples increased as
cycle number increased. This increased range also was larger for
specific HiFi DNA polymerases.

\textbf{Figure 2: Total Number of OTUs in Mock Samples by Number of
Cycles.} The points represent the median number of OTUs for the mock
samples. The lines represent the range of the minimum and maximum number
of OTUs detected within the four fecal samples. The dotted black line
represents the number of OTUs detected when only the references
sequences for the mock community are clustered. The range in the number
of OTUs detected in the mock samples increased as cycle number
increased. This range was also larger for specific HiFi DNA polymerases.

\textbf{Figure 3: Bray-Curtis Community Differences by Five-Cycle
Intervals.} A) within person differences based on the next 5-cycle PCR
interval in fecal samples. B) Within replicate difference based on the
next 5-cycle PCR interval in mock samples. The points represent the
median Bray-Curtis index for the samples. The lines represent the range
of the minimum and maximum Bray-Curits index value for each PCR 5-cycle
increment comparison. The closer a sample is to a Bray-Curtis index of
1.00 the more dissimilar the bacterial community is of the two compared
number of cycles.
